---
title: "Practical Machine Learning Course Project"
author: "Mauro Montauti"
date: "March 22, 2015"
output: html_document
---

## Synopsis

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The goal of this project is to predict the manner of performing unilateral dumbbell biceps curls based on data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. The 5 possible methods include -

A: Exactly according to the specification
B: Throwing the elbows to the front
C: Lifting the dumbbell only halfway
D: Lowering the dumbbell only halfway
E: Throwing the hips to the front

### Setting up and loading data

```{r}
setwd("~/Documents/Machine_learning")

library(caret)

trainingRaw <- read.csv(file="pml-training.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))
testingRaw <- read.csv(file="pml-testing.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))

trainingRaw$classe <- as.factor(trainingRaw$classe)  

```

### Exploring and cleaning the data

```{r}
str(trainingRaw)
NAindex <- apply(trainingRaw,2,function(x) {sum(is.na(x))}) 
trainingRaw <- trainingRaw[,which(NAindex == 0)]
NAindex <- apply(testingRaw,2,function(x) {sum(is.na(x))}) 
testingRaw <- testingRaw[,which(NAindex == 0)]
```

### Pre processing the data 

```{r}
v <- which(lapply(trainingRaw, class) %in% "numeric")

preObj <-preProcess(trainingRaw[,v],method=c('knnImpute', 'center', 'scale'))
trainLess1 <- predict(preObj, trainingRaw[,v])
trainLess1$classe <- trainingRaw$classe

testLess1 <-predict(preObj,testingRaw[,v])
```

### Removing Near Zero variables
As near zero values will not have much meaning in the predictions we'll remove them.

```{r}
nzv <- nearZeroVar(trainLess1,saveMetrics=TRUE)
trainLess1 <- trainLess1[,nzv$nzv==FALSE]

nzv <- nearZeroVar(testLess1,saveMetrics=TRUE)
testLess1 <- testLess1[,nzv$nzv==FALSE]
```


### Cross validation set

The training set is divided in two parts, one for training and the other for cross validation

```{r}
set.seed(12031987)
inTrain = createDataPartition(trainLess1$classe, p = 3/4, list=FALSE)
training = trainLess1[inTrain,]
crossValidation = trainLess1[-inTrain,]
```

### Train Model

The model is created with ramdom forest due to its accuracy based on a set of 28 variables, cross validation will serve as a control method.

```{r}
startTime <- Sys.time();

modFit <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE )

endTime <- Sys.time()
endTime - startTime
```

set.seed(614)
library(lattice); library(ggplot2); library(caret)
pml.training <- read.csv("~/R/Coursera /Machine Learning/Project/pml-training.csv")
inTrain <- createDataPartition(y=pml.training$classe, p=0.9, list=FALSE)
training <- pml.training[inTrain,]
testing <- pml.training[-inTrain,]

### Predictions on the real testing set

```{r}
testingPred <- predict(modFit, testLess1)
testingPred
```

